### 2024-06-11 code_data/plausibility_norming/plausibility_report_HighvsLowSem.Rmd
edits to plausibility code

---
title: "Plausibility Norming"
author: "Pia Schoknecht"
date: "17 06 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lme4)
library(ordinal)
library(brms)
```

# Load and check
```{r load}
# load data
df <- read.csv("norming_data_HighvsLowSem.csv")

head(df)

#dfp_ex <- dfp  %>% filter(subject !="plausib_anim19" & subject !="plausib_anim1")

uniq <- df[!duplicated(df[2]),]
(N_subj <- length(unique(df$subject)))

## SV: isn't this simpler than your obscure code above?
length(unique(df$md5_hash_participant_IP))

colnames(df)

xtabs(~ latin_square_list, data=df)/160
```

# Ratings
```{r}
df$condition <- factor(df$condition, levels = c("c", "d", "f"))

## SV: this calculation is incorrect as it is not aggregated first:
df %>% group_by(condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE))

ggplot(df, aes(x=condition, y=value))+
  geom_jitter(color="lightblue", alpha=0.2)+
  geom_boxplot(fill="grey", alpha=0.7)+
  stat_summary(fun=mean, geom="point", shape=18, size=4) + #adding the mean
  expand_limits(y=c(0.5,7.5))+
  scale_y_continuous(name="plausibility rating", breaks=seq(1, 7, 1))+
  scale_x_discrete(name="", labels=c("c" = "low semantic\ninterference condition", "d" = "high semantic\ninterference condition", "f" = "implausible filler"))+
  theme_bw()+
  labs(title = "Plausibility ratings",
      subtitle = "1 = absolutely implausible; 7 = absolutely plausible")

ggsave("plausibility_ratings.jpg", dpi=600, width=4, height=4)

```

# t-test
```{r}

test <- df %>% filter(condition != "f") %>%
             group_by(subject, condition) %>%
             dplyr::summarize(mean_rating = mean(value))

## SV: redone so I can understand it
xtabs(~subject+condition,df_crit)

df_crit<-subset(df,condition!="f")

xtabs(~subject+condition,df_crit)

xtabs(~item+condition,df_crit)

## c is low, d is high int:
df_crit$cond<-ifelse(df_crit$condition=="c",0.5,-0.5)

library(lme4)
m<-lmer(value~cond + (1+cond|subject) + (1+cond|item),df_crit)
## low semantic int conditions are more plausible than high:
summary(m)
library(car)
## normal approximation seems OK:
qqPlot(residuals(m))

## looks OK:
acf(residuals(m))

## SV: why are we doing this t-test at all?
test_wider <- test %>% pivot_wider(names_from = condition, values_from = mean_rating)

head(test_wider)

## SV: You can see that this t-test is over-enthusiastic because it
## hides significant by-item variance by averaging it away (see the lmer
## results above--by item variability is remarkably high)
t.test(test_wider$c, test_wider$d, paired=TRUE)

```

# Ratings per participant and per item
```{r scale_p}
N <- length(df$value)

p_scale_use <- df %>%
             group_by(subject, condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE), 
                              se_rating = sd_rating / sqrt(N))

ggplot(p_scale_use, aes(x=condition, y=mean_rating, colour=subject)) +
  geom_point(size=1, position = position_dodge(width = 0.3))+
geom_errorbar(aes(ymin=mean_rating-se_rating, ymax=mean_rating+se_rating),
              width=1, position= position_dodge(width = 0.3)) +  
  facet_wrap(.~subject, nrow=4)+
  xlab("condition")+
  ylab("mean_rating")+
  theme_bw()+
  theme(legend.position="none")+
  ggtitle("Pausibility Ratings per subject \n 7 = absolutely plausible, 1 = absolutely implausible \n errorbars = SE")
ggsave("scale_use.jpg", dpi=600, width=5)

#### PER ITEM
p_scale_use <- df %>% filter(condition != "f") %>%
             group_by(item, condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE), 
                              se_rating = sd_rating / sqrt(N))

write.csv(p_scale_use, "plausib_rating_semantics.csv")

p_scale_use$condition <- factor(p_scale_use$condition, levels = c("d", "c"))

items_wide <- p_scale_use %>% select(-sd_rating, -se_rating) %>%
                          pivot_wider(names_from = condition, values_from = mean_rating) %>%
                          mutate(diff = c-d)

ggplot(p_scale_use, aes(x=item, y=mean_rating, colour=condition, group=item)) +
  geom_point(size=1, position = position_dodge(width = 0.3))+
  geom_line(color="grey")+
geom_errorbar(aes(ymin=mean_rating-se_rating, ymax=mean_rating+se_rating),
              width=1, position= position_dodge(width = 0.3)) +  
  xlab("item")+
  ylab("mean_rating")+
  scale_color_discrete(name="", labels=c("c" = "low semantic\ninterference condition", "d" = "high semantic\ninterference condition", "f" = "implausible filler"))+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("Pausibility Ratings per item \n 7 = absolutely plausible, 1 = absolutely implausible \n errorbars = SE")

```

# Does the plausibility affect reading times in the pre-critical region?
```{r}
precrit <- read.csv("../SPR/data/pandora_spr_774_precrit.csv")

precrit_plausib <- inner_join(precrit, p_scale_use, by=c("item", "condition"))

# null model
m_null <- lmer(log(rt) ~ 1 +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

# raw plausibility rating as predictor
library(MASS)
m_lm<-lm(rt~mean_rating,precrit_plausib)

m_plausib <- lmer(log(rt) ~ mean_rating +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)
```

```{r}
summary(m_plausib)
## strong skew:
qqPlot(residuals(m_plausib)
### massive autocorrelation:       
acf(residuals(m_plausib))

m_plausib <- lmer(log(rt) ~ scale(mean_rating,scale=FALSE) +
                       (1 + scale(mean_rating,scale=FALSE)|fullname)+  
                       (1 + scale(mean_rating,scale=FALSE)|item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_plausib)
### massive autocorrelation remains. Why?       
acf(residuals(m_plausib))
```

SV: here, instead of fitting to mean ratings only, can you store the mean by-item ratings as well as the SE in the data frame used for the inner join above? Then center the mean ratings. 
So, the data frame used in the model m_plausib should then have not just the centered mean_rating but also the SE. Then the measurement error model would be (possible in brms only):

```
mbrms<-brm(formula = rt ~ me(c_mean_rating, se_mean_rating) +
(1 | fullname) + (1|item),
data = precrit_plausib, family = lognormal(), prior = some_uninf_priors,
iter = 2000, chains = 4,
control = list(adapt_delta = 0.999,
max_treedepth = 15))
```

SV: Then use the measurement error syntax in the model below where you compare sem with ratings.

# semantic interference as predictor
m_sem <- lmer(log(rt) ~ sem +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_sem)

# semantic interference AND plausibility as predictors
m_sem.plausib <- lmer(log(rt) ~ sem + scale(mean_rating)+
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_sem.plausib)

# do they interact?
m_interaction <- lmer(log(rt) ~ sem*scale(mean_rating)+
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_interaction)
```

# look at reading times for high vs. low plausibility and high vs. low semantic interference

```{r}
# dichotomize plausibility ratings
precrit_plausib$plausibility_cat <- ifelse(precrit_plausib$mean_rating > 5.9, "high",
                                                ifelse(precrit_plausib$mean_rating < 5.5, "low", "med"))

table(precrit_plausib$plausibility_cat)

precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(plausibility_cat) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))
```

```{r}
precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(sem) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))

```

```{r}
precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(sem, plausibility_cat) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))

```