### 2024-07-13 code_data/previous_studies/PreviousFindings.Rmd
Updated via the Open Science Framework

---
title: "Plot previous findings"
author: "Pia Schoknecht"
output: pdf_document
---

```{r}
library(ggplot2)
library(tidyverse)
library(cowplot)
```

# Van Dyke (2007) estimates from published paper:
## E1 
### critical

```{r}
## E1 critical

## not sig. Syn effect in E1 crit:
VD07E1critSynmn <- (875+871)/2 - (912+858)/2
VD07E1critSynSE <- 30
VD07E1critSynlower <- round(VD07E1critSynmn - 2*VD07E1critSynSE)
VD07E1critSynupper <- round(VD07E1critSynmn + 2*VD07E1critSynSE)

## sig. Sem effect in E1 crit:
VD07E1critSemmn <- (875+912)/2 - (871+858)/2
VD07E1critSemF1 <- 6.33
VD07E1critSemSE <- VD07E1critSemmn/sqrt(VD07E1critSemF1)
VD07E1critSemlower <- round(VD07E1critSemmn - 2*VD07E1critSemSE)
VD07E1critSemupper <- round(VD07E1critSemmn + 2*VD07E1critSemSE)

## not sig. Interaction effect in E1 crit:
VD07E1critIntmn <- (875-871) - (912-858)
VD07E1critIntF1 <- 2.36
VD07E1critIntSE <- VD07E1critIntmn/sqrt(VD07E1critIntF1)
VD07E1critIntlower <- round(VD07E1critIntmn - 2*VD07E1critIntSE)
VD07E1critIntupper <- round(VD07E1critIntmn + 2*VD07E1critIntSE)
```

### spill-over
```{r}
## E1 spill-over
## not sig. Syn effect in E1 spill-over:
VD07E1spillSynmn <- (568+551)/2 - (571+566)/2
VD07E1spillSynSE <- 18 #highest SE in spill-over region, Table 3
VD07E1spillSynlower <- round(VD07E1spillSynmn - 2*VD07E1spillSynSE)
VD07E1spillSynupper <- round(VD07E1spillSynmn + 2*VD07E1spillSynSE)

## sig. Sem effect in  E1 spill-over:
VD07E1spillSemmn <- (568+571)/2 - (551+566)/2
VD07E1spillSemF1 <- 18 #highest SE in spill-over region, Table 3
VD07E1spillSemSE <- VD07E1spillSemmn/sqrt(VD07E1spillSemF1)
VD07E1spillSemlower <- round(VD07E1spillSemmn - 2*VD07E1spillSemSE)
VD07E1spillSemupper <- round(VD07E1spillSemmn + 2*VD07E1spillSemSE)

## not sig. Interaction effect in E1 spill-over:
VD07E1spillIntmn <- (568-551) - (571-566)
VD07E1spillIntF1 <- 18 #highest SE in spill-over region, Table 3
VD07E1spillIntSE <- VD07E1spillIntmn/sqrt(VD07E1spillIntF1)
VD07E1spillIntlower <- round(VD07E1spillIntmn - 2*VD07E1spillIntSE)
VD07E1spillIntupper <- round(VD07E1spillIntmn + 2*VD07E1spillIntSE)
```

### Sentence-Final
```{r}
## E1 sentence-final

##  Syn effect in E1 final:
VD07E1finalSynmn <- (822+667)/2 - (697+723)/2
VD07E1finalSyntval <- sqrt(0.99) # F1 < 1 (Table 4)
VD07E1finalSynSE <- VD07E1finalSynmn/VD07E1finalSyntval
VD07E1finalSynlower <- round(VD07E1finalSynmn - 2*VD07E1finalSynSE)
VD07E1finalSynupper <- round(VD07E1finalSynmn + 2*VD07E1finalSynSE)

## Sem effect in  E1 final:
VD07E1finalSemmn <- (822+697)/2 - (667+723)/2
VD07E1finalSemtval <- sqrt(3.06) # F1 (Table 4)
VD07E1finalSemSE <- VD07E1finalSemmn/VD07E1finalSemtval
VD07E1finalSemlower <- round(VD07E1finalSemmn - 2*VD07E1finalSemSE)
VD07E1finalSemupper <- round(VD07E1finalSemmn + 2*VD07E1finalSemSE)

## Interaction effect in E1 final:
VD07E1finalIntmn <- (822-667) - (697-723)
VD07E1finalInttval <- sqrt(5.39) # F1 (Table 4)
VD07E1finalIntSE <- VD07E1finalIntmn/VD07E1finalInttval
VD07E1finalIntlower <- round(VD07E1finalIntmn - 2*VD07E1finalIntSE)
VD07E1finalIntupper <- round(VD07E1finalIntmn + 2*VD07E1finalIntSE)
```

## Experiment 2
### critical: SYN 
```{r}
# Experiment 2 SYN in critical
## sig. Syn effect in E2 crit region FPRT:
VD07E2critSynFPRTmn <- (413+418)/2-(376+382)/2
VD07E2critSynFPRTtval <- sqrt(8.32)
VD07E2critSynFPRTSE <- VD07E2critSynFPRTmn/VD07E2critSynFPRTtval
VD07E2critSynFPRTlower <- round(VD07E2critSynFPRTmn - 2*VD07E2critSynFPRTSE)
VD07E2critSynFPRTupper <- round(VD07E2critSynFPRTmn + 2*VD07E2critSynFPRTSE)

# sig. Syn effect in E2 crit region RPD:
VD07E2critSynRPDmn <- (594+663)/2-(454+495)/2
VD07E2critSynRPDtval <- sqrt(27.21)
VD07E2critSynRPDSE <- VD07E2critSynRPDmn/VD07E2critSynRPDtval
VD07E2critSynRPDlower <- round(VD07E2critSynRPDmn - 2*VD07E2critSynRPDSE)
VD07E2critSynRPDupper <- round(VD07E2critSynRPDmn + 2*VD07E2critSynRPDSE)

## sig. Syn effect in E2 crit region TFT:
VD07E2critSynTFTmn <- (738+761)/2-(630+653)/2
VD07E2critSynTFTtval <- sqrt(29.95)
VD07E2critSynTFTSE <- VD07E2critSynTFTmn/VD07E2critSynTFTtval
VD07E2critSynTFTlower <- round(VD07E2critSynTFTmn - 2*VD07E2critSynTFTSE)
VD07E2critSynTFTupper <- round(VD07E2critSynTFTmn + 2*VD07E2critSynTFTSE)
```

### critical: SEM
```{r}
# Experiment 2 Sem in critical
## non sig. Sem effect in E2 crit region FPRT:
VD07E2critSemFPRTmn <- (418+382)/2-(413+376)/2
VD07E2critSemFPRTtval <- sqrt(1.13)
VD07E2critSemFPRTSE <- VD07E2critSemFPRTmn/VD07E2critSemFPRTtval
VD07E2critSemFPRTlower <- round(VD07E2critSemFPRTmn - 2*VD07E2critSemFPRTSE)
VD07E2critSemFPRTupper <- round(VD07E2critSemFPRTmn + 2*VD07E2critSemFPRTSE)

# non sig. Sem effect in E2 crit region RPD:
VD07E2critSemRPDmn <- (663+495)/2-(594+454)/2
VD07E2critSemRPDtval <- sqrt(3.24)
VD07E2critSemRPDSE <- VD07E2critSemRPDmn/VD07E2critSemRPDtval
VD07E2critSemRPDlower <- round(VD07E2critSemRPDmn - 2*VD07E2critSemRPDSE)
VD07E2critSemRPDupper <- round(VD07E2critSemRPDmn + 2*VD07E2critSemRPDSE)

## non sig. Sem effect in E2 crit region TFT:
VD07E2critSemTFTmn <- (761+653)/2-(738+630)/2
VD07E2critSemTFTSE <- 42 #highest SE in critical region, Table 7
VD07E2critSemTFTlower <- round(VD07E2critSemTFTmn - 2*VD07E2critSemTFTSE)
VD07E2critSemTFTupper <- round(VD07E2critSemTFTmn + 2*VD07E2critSemTFTSE)
```

### critical: Int
```{r}
# Experiment 2 Int in critical
## non sig. Int effect in E2 crit region FPRT:
VD07E2critIntFPRTmn <- (418-413)-(382-376)
VD07E2critIntFPRTSE <- 21 #highest SE in critical region, Table 7
VD07E2critIntFPRTlower <- round(VD07E2critIntFPRTmn - 2*VD07E2critIntFPRTSE)
VD07E2critIntFPRTupper <- round(VD07E2critIntFPRTmn + 2*VD07E2critIntFPRTSE)

# non sig. Int effect in E2 crit region RPD:
VD07E2critIntRPDmn <- (663-594)-(495-454)
VD07E2critIntRPDSE <- 44 #highest SE in critical region, Table 7
VD07E2critIntRPDlower <- round(VD07E2critIntRPDmn - 2*VD07E2critIntRPDSE)
VD07E2critIntRPDupper <- round(VD07E2critIntRPDmn + 2*VD07E2critIntRPDSE)

## non sig. Int effect in E2 crit region TFT:
VD07E2critIntTFTmn <- (761-738)-(653-630)
VD07E2critIntTFTSE <- 42 #highest SE in critical region, Table 7
VD07E2critIntTFTlower <- round(VD07E2critIntTFTmn - 2*VD07E2critIntTFTSE)
VD07E2critIntTFTupper <- round(VD07E2critIntTFTmn + 2*VD07E2critIntTFTSE)
```

### spill-over: SYN 
```{r}
# Experiment 2 SYN in spill-over
## almost sig. Syn effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritSynFPRTmn<-(325+296)/2-(320+364)/2
VD07E2postcritSynFPRTtval<-sqrt(6.58)
VD07E2postcritSynFPRTSE<-VD07E2postcritSynFPRTmn/VD07E2postcritSynFPRTtval
VD07E2postcritSynFPRTlower<-VD07E2postcritSynFPRTmn-2*VD07E2postcritSynFPRTSE
VD07E2postcritSynFPRTupper<-VD07E2postcritSynFPRTmn+2*VD07E2postcritSynFPRTSE

## almost sig. Syn effect in E2 post-crit region RPD:
VD07E2postcritSynRPDmn <- (1365+1295)/2-(970+1205)/2
VD07E2postcritSynRPDtval <- sqrt(12.17)
VD07E2postcritSynRPDSE <- VD07E2postcritSynRPDmn/VD07E2postcritSynRPDtval
VD07E2postcritSynRPDlower <- round(VD07E2postcritSynRPDmn - 2*VD07E2postcritSynRPDSE)
VD07E2postcritSynRPDupper <- round(VD07E2postcritSynRPDmn + 2*VD07E2postcritSynRPDSE)

## non sig. Syn effect in E2 post-crit region TFT:
VD07E2postcritSynTFTmn <- (491+493)/2-(502+540)/2
VD07E2postcritSynTFTSE <- 30 #highest SE in spill-over region, Table 7
VD07E2postcritSynTFTlower <- round(VD07E2postcritSynTFTmn - 2*VD07E2postcritSynTFTSE)
VD07E2postcritSynTFTupper <- round(VD07E2postcritSynTFTmn + 2*VD07E2postcritSynTFTSE)
```

### spill-over: Sem 
```{r}
# Experiment 2 Sem in spill-over
## almost sig. Sem effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritSemFPRTmn<- (296+364)/2-(325+320)/2
VD07E2postcritSemFPRTSE<- 21 #highest SE in spill-over region, Table 7
VD07E2postcritSemFPRTlower<-VD07E2postcritSemFPRTmn-2*VD07E2postcritSemFPRTSE
VD07E2postcritSemFPRTupper<-VD07E2postcritSemFPRTmn+2*VD07E2postcritSemFPRTSE

## almost sig. Sem effect in E2 post-crit region RPD:
VD07E2postcritSemRPDmn <- (1295+1205)/2-(1365+970)/2
VD07E2postcritSemRPDSE <- 140 #highest SE in spill-over region, Table 7
VD07E2postcritSemRPDlower <- round(VD07E2postcritSemRPDmn - 2*VD07E2postcritSemRPDSE)
VD07E2postcritSemRPDupper <- round(VD07E2postcritSemRPDmn + 2*VD07E2postcritSemRPDSE)

## almost sig. Sem effect in E2 post-crit region TFT:
VD07E2postcritSemTFTmn <- (493+540)/2-(491+502)/2
VD07E2postcritSemTFTSE <-  30 #highest SE in spill-over region, Table 7
VD07E2postcritSemTFTlower <- round(VD07E2postcritSemTFTmn - 2*VD07E2postcritSemTFTSE)
VD07E2postcritSemTFTupper <- round(VD07E2postcritSemTFTmn + 2*VD07E2postcritSemTFTSE)
```

### spill-over: Int 
```{r}
# Experiment 2 Int in spill-over
## almost sig. Int effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritIntFPRTmn<- (296-325)-(364-320)
VD07E2postcritIntFPRTtval<- sqrt(5.22)
VD07E2postcritIntFPRTSE<-VD07E2postcritIntFPRTmn/VD07E2postcritIntFPRTtval
VD07E2postcritIntFPRTlower<-VD07E2postcritIntFPRTmn-2*VD07E2postcritIntFPRTSE
VD07E2postcritIntFPRTupper<-VD07E2postcritIntFPRTmn+2*VD07E2postcritIntFPRTSE

## almost sig. Int effect in E2 post-crit region RPD:
VD07E2postcritIntRPDmn <- (1295-1365)-(1205-970)
VD07E2postcritIntRPDtval <- sqrt(3.45)
VD07E2postcritIntRPDSE <- VD07E2postcritIntRPDmn/VD07E2postcritIntRPDtval
VD07E2postcritIntRPDlower <- round(VD07E2postcritIntRPDmn - 2*VD07E2postcritIntRPDSE)
VD07E2postcritIntRPDupper <- round(VD07E2postcritIntRPDmn + 2*VD07E2postcritIntRPDSE)

## almost sig. Int effect in E2 post-crit region TFT:
VD07E2postcritIntTFTmn <- (493-491)-(540-502)
VD07E2postcritIntTFTSE <- 30 #highest SE in spill-over region, Table 7
VD07E2postcritIntTFTlower <- round(VD07E2postcritIntTFTmn - 2*VD07E2postcritIntTFTSE)
VD07E2postcritIntTFTupper <- round(VD07E2postcritIntTFTmn + 2*VD07E2postcritIntTFTSE)
```

### Sentence-Final: SYN 
```{r}
# Experiment 2 SYN in sentence-final region (Table 7)
## Syn effect in E2 sentence-final region FPRT 
VD07E2finalSynFPRTmn<-(259+271)/2-(286+274)/2
VD07E2finalSynFPRTSE<- 22 #highest SE in sentence-final region, Table 7
VD07E2finalSynFPRTlower<-VD07E2finalSynFPRTmn-2*VD07E2finalSynFPRTSE
VD07E2finalSynFPRTupper<-VD07E2finalSynFPRTmn+2*VD07E2finalSynFPRTSE

## Syn effect in E2 sentence-final region RPD:
VD07E2finalSynRPDmn <- (1925+2131)/2-(1695+1806)/2 
VD07E2finalSynRPDtval <- sqrt(4.74) # F1 Table 10
VD07E2finalSynRPDSE <- VD07E2finalSynRPDmn/VD07E2finalSynRPDtval
VD07E2finalSynRPDlower <- round(VD07E2finalSynRPDmn - 2*VD07E2finalSynRPDSE)
VD07E2finalSynRPDupper <- round(VD07E2finalSynRPDmn + 2*VD07E2finalSynRPDSE)

## Syn effect in E2 sentence-final region TFT:
VD07E2finalSynTFTmn <- (349+360)/2-(362+373)/2
VD07E2finalSynTFTSE<- 40 #highest SE in sentence-final region, Table 7
VD07E2finalSynTFTlower <- round(VD07E2finalSynTFTmn - 2*VD07E2finalSynTFTSE)
VD07E2finalSynTFTupper <- round(VD07E2finalSynTFTmn + 2*VD07E2finalSynTFTSE)
```

### sentence-final: Sem 
```{r}
# Experiment 2 Sem in sentence-final (Table 7)
## Sem effect in E2 sentence-final region FPRT
VD07E2finalSemFPRTmn<- (271+274)/2-(259+286)/2
VD07E2finalSemFPRTSE<- 22 #highest SE in sentence-final region, Table 7
VD07E2finalSemFPRTlower<-VD07E2finalSemFPRTmn-2*VD07E2finalSemFPRTSE
VD07E2finalSemFPRTupper<-VD07E2finalSemFPRTmn+2*VD07E2finalSemFPRTSE

## Sem effect in E2 sentence-final region RPD:
VD07E2finalSemRPDmn <- (2131+1806)/2-(1925+1695)/2
VD07E2finalSemRPDtval <- sqrt(1.69) # F1 Table 10
VD07E2finalSemRPDSE <- VD07E2finalSemRPDmn/VD07E2finalSemRPDtval
VD07E2finalSemRPDlower <- round(VD07E2finalSemRPDmn - 2*VD07E2finalSemRPDSE)
VD07E2finalSemRPDupper <- round(VD07E2finalSemRPDmn + 2*VD07E2finalSemRPDSE)

## Sem effect in E2 sentence-final region TFT:
VD07E2finalSemTFTmn <- (360+373)/2-(349+362)/2
VD07E2finalSemTFTSE <- 40 #highest SE in sentence-final region, Table 7
VD07E2finalSemTFTlower <- round(VD07E2finalSemTFTmn - 2*VD07E2finalSemTFTSE)
VD07E2finalSemTFTupper <- round(VD07E2finalSemTFTmn + 2*VD07E2finalSemTFTSE)
```

### sentence-final: Int 
```{r}
# Experiment 2 Int in sentence-final (Table 7)
## Int effect in E2 sentence-final region FPRT
VD07E2finalIntFPRTmn<- (271-259)-(274-286)
VD07E2finalIntFPRTSE<- 22 #highest SE in sentence-final region, Table 7
VD07E2finalIntFPRTlower<-VD07E2finalIntFPRTmn-2*VD07E2finalIntFPRTSE
VD07E2finalIntFPRTupper<-VD07E2finalIntFPRTmn+2*VD07E2finalIntFPRTSE

## Int effect in E2 sentence-final region RPD:
VD07E2finalIntRPDmn <- (2131-1925)-(1806-1695)
VD07E2finalIntRPDtval <- sqrt(0.99) # F1 < 1, Table 10
VD07E2finalIntRPDSE <- VD07E2finalIntRPDmn/VD07E2finalIntRPDtval
VD07E2finalIntRPDlower <- round(VD07E2finalIntRPDmn - 2*VD07E2finalIntRPDSE)
VD07E2finalIntRPDupper <- round(VD07E2finalIntRPDmn + 2*VD07E2finalIntRPDSE)

## Int effect in E2 sentence-final region TFT:
VD07E2finalIntTFTmn <- (360-349)-(373-362)
VD07E2finalIntTFTSE <-  40 #highest SE in sentence-final region, Table 7
VD07E2finalIntTFTlower <- round(VD07E2finalIntTFTmn - 2*VD07E2finalIntTFTSE)
VD07E2finalIntTFTupper <- round(VD07E2finalIntTFTmn + 2*VD07E2finalIntTFTSE)
```

## Experiment 3  
### precrit: SYN
```{r}
# Experiment 3  SYN in pre-crit
## almost sig. Syn effect in E3 precrit region FPRT:
VD07E3precritSynFPRTmn <- (462+447)/2-(449+455)/2
VD07E3precritSynFPRTSE <- 18 #highest SE in pre-crit region, Table 16
VD07E3precritSynFPRTlower <- round(VD07E3precritSynFPRTmn-2*VD07E3precritSynFPRTSE)
VD07E3precritSynFPRTupper <- round(VD07E3precritSynFPRTmn+2*VD07E3precritSynFPRTSE)

## almost sig. Syn effect in E3 precrit region RPD:
VD07E3precritSynRPDmn <- (490+551)/2-(472+493)/2
VD07E3precritSynRPDtval <- sqrt(6.59)
VD07E3precritSynRPDSE <- VD07E3precritSynRPDmn/VD07E3precritSynRPDtval
VD07E3precritSynRPDlower <- round(VD07E3precritSynRPDmn-2*VD07E3precritSynRPDSE)
VD07E3precritSynRPDupper <- round(VD07E3precritSynRPDmn+2*VD07E3precritSynRPDSE)

## almost sig. effect in E3 precrit region TFT:
VD07E3precritSynTFTmn <- (622+710)/2-(586+640)/2
VD07E3precritSynTFTtval <- sqrt(7.29)
VD07E3precritSynTFTSE <- VD07E3precritSynTFTmn/VD07E3precritSynTFTtval
VD07E3precritSynTFTlower <- round(VD07E3precritSynTFTmn-2*VD07E3precritSynTFTSE)
VD07E3precritSynTFTupper <- round(VD07E3precritSynTFTmn+2*VD07E3precritSynTFTSE)
```

### precrit: SEM
```{r}
# Experiment 3  SEM in pre-crit
## almost sig. Sem effect in E3 precrit region FPRT:
VD07E3precritSemFPRTmn <- (455+447)/2-(449+462)/2
VD07E3precritSemFPRTSE <- 18 #highest SE in pre-crit region, Table 16
VD07E3precritSemFPRTlower <- round(VD07E3precritSemFPRTmn-2*VD07E3precritSemFPRTSE)
VD07E3precritSemFPRTupper <- round(VD07E3precritSemFPRTmn+2*VD07E3precritSemFPRTSE)

## almost sig. Sem effect in E3 precrit region RPD:
VD07E3precritSemRPDmn <- (493+551)/2-(472+490)/2
VD07E3precritSemRPDtval <- sqrt(4.89)
VD07E3precritSemRPDSE <- VD07E3precritSemRPDmn/VD07E3precritSemRPDtval
VD07E3precritSemRPDlower <- round(VD07E3precritSemRPDmn-2*VD07E3precritSemRPDSE)
VD07E3precritSemRPDupper <- round(VD07E3precritSemRPDmn+2*VD07E3precritSemRPDSE)

## almost sig. Sem effect in E3 precrit region TFT:
VD07E3precritSemTFTmn <- (640+710)/2-(586+622)/2
VD07E3precritSemTFTtval <- sqrt(10.73)
VD07E3precritSemTFTSE <-  VD07E3precritSemTFTmn/VD07E3precritSemTFTtval
VD07E3precritSemTFTlower <- round(VD07E3precritSemTFTmn-2*VD07E3precritSemTFTSE)
VD07E3precritSemTFTupper <- round(VD07E3precritSemTFTmn+2*VD07E3precritSemTFTSE)
```

### precrit: INT
```{r}
# Experiment 3  Int in pre-crit
## almost sig. Int effect in E3 precrit region FPRT:
VD07E3precritIntFPRTmn <- (447-462)-(455-449)
VD07E3precritIntFPRTSE <- 18 #highest SE in pre-crit region, Table 16
VD07E3precritIntFPRTlower <- round(VD07E3precritIntFPRTmn-2*VD07E3precritIntFPRTSE)
VD07E3precritIntFPRTupper <- round(VD07E3precritIntFPRTmn+2*VD07E3precritIntFPRTSE)

## almost sig. Int effect in E3 precrit region RPD:
VD07E3precritIntRPDmn <- (551-490)-(493-472)
VD07E3precritIntRPDSE <- 36 #highest SE in pre-crit region, Table 16
VD07E3precritIntRPDlower <- round(VD07E3precritIntRPDmn-2*VD07E3precritIntRPDSE)
VD07E3precritIntRPDupper <- round(VD07E3precritIntRPDmn+2*VD07E3precritIntRPDSE)

## almost sig. Int effect in E3 precrit region TFT:
VD07E3precritIntTFTmn <- (710-622)-(640-586)
VD07E3precritIntTFTSE <- 47 #highest SE in pre-crit region, Table 16
VD07E3precritIntTFTlower <- round(VD07E3precritIntTFTmn-2*VD07E3precritIntTFTSE)
VD07E3precritIntTFTupper <- round(VD07E3precritIntTFTmn+2*VD07E3precritIntTFTSE)
```

### crit: SYN
```{r}
# Experiment 3  SYN in crit

## almost sig. Syn effect in E3 crit region FPRT:
VD07E3critSynFPRTmn <- (294+280)/2-(274+282)/2
VD07E3critSynFPRTtval <- sqrt(3.95) # reported as <3.96
VD07E3critSynFPRTSE <- VD07E3critSynFPRTmn/VD07E3critSynFPRTtval
VD07E3critSynFPRTlower <- round(VD07E3critSynFPRTmn-2*VD07E3critSynFPRTSE)
VD07E3critSynFPRTupper <- round(VD07E3critSynFPRTmn+2*VD07E3critSynFPRTSE)

## almost sig. Syn effect in E3 crit region RPD:
VD07E3critSynRPDmn <- (354+344)/2-(314+315)/2
VD07E3critSynRPDtval <- sqrt(6.15)
VD07E3critSynRPDSE <- VD07E3critSynRPDmn/VD07E3critSynRPDtval
VD07E3critSynRPDlower <- round(VD07E3critSynRPDmn-2*VD07E3critSynRPDSE)
VD07E3critSynRPDupper <- round(VD07E3critSynRPDmn+2*VD07E3critSynRPDSE)

## almost sig. Syn effect in E3 crit region TFT:
VD07E3critSynTFTmn <- (451+467)/2-(414+421)/2
VD07E3critSynTFTSE <- 38 #highest SE in crit region, Table 16
VD07E3critSynTFTlower <- round(VD07E3critSynTFTmn-2*VD07E3critSynTFTSE)
VD07E3critSynTFTupper <- round(VD07E3critSynTFTmn+2*VD07E3critSynTFTSE)
```

### crit: Sem 
```{r}
# Experiment 3  SEM in crit
## almost sig. Sem effect in E3 crit region FPRT:
VD07E3critSemFPRTmn <- (282+280)/2-(274+294)/2
VD07E3critSemFPRTSE <- 11 #highest SE in crit region, Table 16
VD07E3critSemFPRTlower <- round(VD07E3critSemFPRTmn-2*VD07E3critSemFPRTSE)
VD07E3critSemFPRTupper <- round(VD07E3critSemFPRTmn+2*VD07E3critSemFPRTSE)

## almost sig. Sem effect in E3 crit region RPD:
VD07E3critSemRPDmn <- (315+344)/2-(314+354)/2
VD07E3critSemRPDSE <- 23 #highest SE in crit region, Table 16
VD07E3critSemRPDlower <- round(VD07E3critSemRPDmn-2*VD07E3critSemRPDSE)
VD07E3critSemRPDupper <- round(VD07E3critSemRPDmn+2*VD07E3critSemRPDSE)

## almost sig. Sem effect in E3 crit region TFT:
VD07E3critSemTFTmn <- (421+467)/2-(414+451)/2
VD07E3critSemTFTSE <- 38 #highest SE in crit region, Table 16
VD07E3critSemTFTlower <- round(VD07E3critSemTFTmn-2*VD07E3critSemTFTSE)
VD07E3critSemTFTupper <- round(VD07E3critSemTFTmn+2*VD07E3critSemTFTSE)
```

### crit: Int 
```{r}
# Experiment 3  Int in crit
## almost sig. Int effect in E3 crit region FPRT:
VD07E3critIntFPRTmn <- (280-294)-(282-274)
VD07E3critIntFPRTtval <- sqrt(3.34) # reported in paper as <3.35
VD07E3critIntFPRTSE <- VD07E3critIntFPRTmn/VD07E3critIntFPRTtval
VD07E3critIntFPRTlower <- round(VD07E3critIntFPRTmn-2*VD07E3critIntFPRTSE)
VD07E3critIntFPRTupper <- round(VD07E3critIntFPRTmn+2*VD07E3critIntFPRTSE)

## almost sig. Int effect in E3 crit region RPD:
VD07E3critIntRPDmn <- (344-354)-(315-314)
VD07E3critIntRPDSE <- 23 #highest SE in crit region, Table 16
VD07E3critIntRPDlower <- round(VD07E3critIntRPDmn-2*VD07E3critIntRPDSE)
VD07E3critIntRPDupper <- round(VD07E3critIntRPDmn+2*VD07E3critIntRPDSE)

## almost sig. Int effect in E3 crit region TFT:
VD07E3critIntTFTmn <- (467-451)-(421-414)
VD07E3critIntTFTSE <- 38 #highest SE in crit region, Table 16
VD07E3critIntTFTlower <- round(VD07E3critIntTFTmn-2*VD07E3critIntTFTSE)
VD07E3critIntTFTupper <- round(VD07E3critIntTFTmn+2*VD07E3critIntTFTSE)
```

### spill-over: SYN
```{r}
# Experiment 3  SYN in postcrit
## almost sig. Syn effect in E3 postcrit region FPRT:
VD07E3postcritSynFPRTmn <- (424+442)/2-(437+448)/2
VD07E3postcritSynFPRTSE <- 24 #highest SE in spill-over region, Table 16
VD07E3postcritSynFPRTlower <- round(VD07E3postcritSynFPRTmn-2*VD07E3postcritSynFPRTSE)
VD07E3postcritSynFPRTupper <- round(VD07E3postcritSynFPRTmn+2*VD07E3postcritSynFPRTSE)

## almost sig. Syn effect in E3 postcrit region RPD:
VD07E3postcritSynRPDmn <- (615+625)/2-(553+607)/2
VD07E3postcritSynRPDtval <- sqrt(2.78) # reported as <2.79
VD07E3postcritSynRPDSE <- VD07E3postcritSynRPDmn/VD07E3postcritSynRPDtval
VD07E3postcritSynRPDlower <- round(VD07E3postcritSynRPDmn-2*VD07E3postcritSynRPDSE)
VD07E3postcritSynRPDupper <- round(VD07E3postcritSynRPDmn+2*VD07E3postcritSynRPDSE)

## almost sig. Syn effect in E3 postcrit region TFT:
VD07E3postcritSynTFTmn <- (766+731)/2-(707+711)/2
VD07E3postcritSynTFTSE <- 59 #highest SE in spill-over region, Table 16
VD07E3postcritSynTFTlower <- round(VD07E3postcritSynTFTmn-2*VD07E3postcritSynTFTSE)
VD07E3postcritSynTFTupper <- round(VD07E3postcritSynTFTmn+2*VD07E3postcritSynTFTSE)
```

### spill-over: Sem 
```{r}
# Experiment 3  SEM in postcrit
## almost sig. Sem effect in E3 postcrit region FPRT:
VD07E3postcritSemFPRTmn <- (437+424)/2-(448+442)/2
VD07E3postcritSemFPRTSE <- 24 #highest SE in spill-over region, Table 16
VD07E3postcritSemFPRTlower <- round(VD07E3postcritSemFPRTmn-2*VD07E3postcritSemFPRTSE)
VD07E3postcritSemFPRTupper <- round(VD07E3postcritSemFPRTmn+2*VD07E3postcritSemFPRTSE)

## almost sig. Sem effect in E3 postcrit region RPD:
VD07E3postcritSemRPDmn <- (553+615)/2-(607+625)/2
VD07E3postcritSemRPDtval <- sqrt(2.78) # reported as <2.79
VD07E3postcritSemRPDSE <- VD07E3postcritSemRPDmn/VD07E3postcritSemRPDtval
VD07E3postcritSemRPDlower <- round(VD07E3postcritSemRPDmn-2*VD07E3postcritSemRPDSE)
VD07E3postcritSemRPDupper <- round(VD07E3postcritSemRPDmn+2*VD07E3postcritSemRPDSE)

## almost sig. Sem effect in E3 postcrit region TFT:
VD07E3postcritSemTFTmn <- (707+766)/2-(711+731)/2
VD07E3postcritSemTFTtval <- sqrt(6.15)
VD07E3postcritSemTFTSE <- VD07E3postcritSemTFTmn/VD07E3postcritSemTFTtval
VD07E3postcritSemTFTlower <- round(VD07E3postcritSemTFTmn-2*VD07E3postcritSemTFTSE)
VD07E3postcritSemTFTupper <- round(VD07E3postcritSemTFTmn+2*VD07E3postcritSemTFTSE)
```

### spill-over: Int 
```{r}
# Experiment 3  Int in postcrit
## almost sig. Int effect in E3 postcrit region FPRT:
VD07E3postcritIntFPRTmn <- (424-442)-(437-448)
VD07E3postcritIntFPRTSE <-24 #highest SE in spill-over region, Table 16
VD07E3postcritIntFPRTlower <- round(VD07E3postcritIntFPRTmn-2*VD07E3postcritIntFPRTSE)
VD07E3postcritIntFPRTupper <- round(VD07E3postcritIntFPRTmn+2*VD07E3postcritIntFPRTSE)

## almost sig. Int effect in E3 postcrit region RPD:
VD07E3postcritIntRPDmn <- (615-625)-(553-607)
VD07E3postcritIntRPDSE <- 41 #highest SE in spill-over region, Table 16
VD07E3postcritIntRPDlower <- round(VD07E3postcritIntRPDmn-2*VD07E3postcritIntRPDSE)
VD07E3postcritIntRPDupper <- round(VD07E3postcritIntRPDmn+2*VD07E3postcritIntRPDSE)

## almost sig. Int effect in E3 postcrit region TFT:
VD07E3postcritIntTFTmn <- (766-731)-(707-711)
VD07E3postcritIntTFTSE <- 59 #highest SE in spill-over region, Table 16
VD07E3postcritIntTFTlower <- round(VD07E3postcritIntTFTmn-2*VD07E3postcritIntTFTSE)
VD07E3postcritIntTFTupper <- round(VD07E3postcritIntTFTmn+2*VD07E3postcritIntTFTSE)
```

### sentence final: SYN
```{r}
# Experiment 3  SYN in sentence final region (Table 16 in VD07)
## almost sig. Syn effect in E3 final region FPRT:
VD07E3finalSynFPRTmn <- (463+483)/2-(480+464)/2
VD07E3finalSynFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalSynFPRTSE <- VD07E3finalSynFPRTmn/VD07E3finalSynFPRTtval
VD07E3finalSynFPRTlower <- round(VD07E3finalSynFPRTmn-2*VD07E3finalSynFPRTSE)
VD07E3finalSynFPRTupper <- round(VD07E3finalSynFPRTmn+2*VD07E3finalSynFPRTSE)

## almost sig. Syn effect in E3 final region RPD:
VD07E3finalSynRPDmn <- (2474+2068)/2-(2147+1875)/2
VD07E3finalSynRPDtval <- sqrt(2.79) # F1
VD07E3finalSynRPDSE <- VD07E3finalSynRPDmn/VD07E3finalSynRPDtval
VD07E3finalSynRPDlower <- round(VD07E3finalSynRPDmn-2*VD07E3finalSynRPDSE)
VD07E3finalSynRPDupper <- round(VD07E3finalSynRPDmn+2*VD07E3finalSynRPDSE)

## almost sig. Syn effect in E3 final region TFT:
VD07E3finalSynTFTmn <- (663+675)/2-(705+649)/2
VD07E3finalSynTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalSynTFTSE <- VD07E3finalSynTFTmn/VD07E3finalSynTFTtval
VD07E3finalSynTFTlower <- round(VD07E3finalSynTFTmn-2*VD07E3finalSynTFTSE)
VD07E3finalSynTFTupper <- round(VD07E3finalSynTFTmn+2*VD07E3finalSynTFTSE)
```

### Sentence-final: Sem 
```{r}
# Experiment 3  SEM in final (Table 16 in VD07)

## Sem effect in E3 final region FPRT:
VD07E3finalSemFPRTmn <- (480+463)/2-(464+483)/2
VD07E3finalSemFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalSemFPRTSE <- VD07E3finalSemFPRTmn/VD07E3finalSemFPRTtval
VD07E3finalSemFPRTlower <- round(VD07E3finalSemFPRTmn-2*VD07E3finalSemFPRTSE)
VD07E3finalSemFPRTupper <- round(VD07E3finalSemFPRTmn+2*VD07E3finalSemFPRTSE)

## Sem effect in E3 final region RPD:
VD07E3finalSemRPDmn <- (2147+2474)/2-(1875+2068)/2
VD07E3finalSemRPDtval <- sqrt(9.26) # F1 (Table 19)
VD07E3finalSemRPDSE <- VD07E3finalSemRPDmn/VD07E3finalSemRPDtval
VD07E3finalSemRPDlower <- round(VD07E3finalSemRPDmn-2*VD07E3finalSemRPDSE)
VD07E3finalSemRPDupper <- round(VD07E3finalSemRPDmn+2*VD07E3finalSemRPDSE)

## Sem effect in E3 final region TFT:
VD07E3finalSemTFTmn <- (705+663)/2-(649+675)/2
VD07E3finalSemTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalSemTFTSE <- VD07E3finalSemTFTmn/VD07E3finalSemTFTtval
VD07E3finalSemTFTlower <- round(VD07E3finalSemTFTmn-2*VD07E3finalSemTFTSE)
VD07E3finalSemTFTupper <- round(VD07E3finalSemTFTmn+2*VD07E3finalSemTFTSE)
```

### Sentence-Final: Int 
```{r}
# Experiment 3  Int in final (Table 16 in VD07)
## Int effect in E3 final region FPRT:
VD07E3finalIntFPRTmn <- (463-483)-(480-464)
VD07E3finalIntFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalIntFPRTSE <- VD07E3finalIntFPRTmn/VD07E3finalIntFPRTtval
VD07E3finalIntFPRTlower <- round(VD07E3finalIntFPRTmn-2*VD07E3finalIntFPRTSE)
VD07E3finalIntFPRTupper <- round(VD07E3finalIntFPRTmn+2*VD07E3finalIntFPRTSE)

## Int effect in E3 final region RPD:
VD07E3finalIntRPDmn <- (2474-2068)-(2147-1875)
VD07E3finalIntRPDtval <- sqrt(0.99) # reported in paper as below 1
VD07E3finalIntRPDSE <- VD07E3finalIntRPDmn/VD07E3finalIntRPDtval
VD07E3finalIntRPDlower <- round(VD07E3finalIntRPDmn-2*VD07E3finalIntRPDSE)
VD07E3finalIntRPDupper <- round(VD07E3finalIntRPDmn+2*VD07E3finalIntRPDSE)

## Int effect in E3 final region TFT:
VD07E3finalIntTFTmn <- (663-675)-(705-649)
VD07E3finalIntTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalIntTFTSE <- VD07E3finalIntTFTmn/VD07E3finalIntTFTtval
VD07E3finalIntTFTlower <- round(VD07E3finalIntTFTmn-2*VD07E3finalIntTFTSE)
VD07E3finalIntTFTupper <- round(VD07E3finalIntTFTmn+2*VD07E3finalIntTFTSE)
```

# Prepare data frame
```{r}
VD07summary<-rbind(c(VD07E1critSynmn, VD07E1critSynlower, VD07E1critSynupper),
                   c(VD07E1critSemmn, VD07E1critSemlower, VD07E1critSemupper),
                   c(VD07E1critIntmn, VD07E1critIntlower, VD07E1critIntupper),

                   c(VD07E1spillSynmn, VD07E1spillSynlower, VD07E1spillSynupper),
                   c(VD07E1spillSemmn, VD07E1spillSemlower, VD07E1spillSemupper),
                   c(VD07E1spillIntmn, VD07E1spillIntlower, VD07E1spillIntupper),

                   c(VD07E1finalSynmn, VD07E1finalSynlower, VD07E1finalSynupper),
                   c(VD07E1finalSemmn, VD07E1finalSemlower, VD07E1finalSemupper),
                   c(VD07E1finalIntmn, VD07E1finalIntlower, VD07E1finalIntupper),

                   c(VD07E2critSynFPRTmn, VD07E2critSynFPRTlower, VD07E2critSynFPRTupper),
                   c(VD07E2critSynRPDmn, VD07E2critSynRPDlower, VD07E2critSynRPDupper),
                   c(VD07E2critSynTFTmn, VD07E2critSynTFTlower, VD07E2critSynTFTupper),

                   c(VD07E2critSemFPRTmn, VD07E2critSemFPRTlower, VD07E2critSemFPRTupper),
                   c(VD07E2critSemRPDmn, VD07E2critSemRPDlower, VD07E2critSemRPDupper),
                   c(VD07E2critSemTFTmn, VD07E2critSemTFTlower, VD07E2critSemTFTupper),

                   c(VD07E2critIntFPRTmn, VD07E2critIntFPRTlower, VD07E2critIntFPRTupper),
                   c(VD07E2critIntRPDmn, VD07E2critIntRPDlower, VD07E2critIntRPDupper),
                   c(VD07E2critIntTFTmn, VD07E2critIntTFTlower, VD07E2critIntTFTupper),

                   c(VD07E2postcritSynFPRTmn, VD07E2postcritSynFPRTlower, VD07E2postcritSynFPRTupper),
                   c(VD07E2postcritSynRPDmn, VD07E2postcritSynRPDlower, VD07E2postcritSynRPDupper),
                   c(VD07E2postcritSynTFTmn, VD07E2postcritSynTFTlower, VD07E2postcritSynTFTupper),

                   c(VD07E2postcritSemFPRTmn, VD07E2postcritSemFPRTlower, VD07E2postcritSemFPRTupper),
                   c(VD07E2postcritSemRPDmn, VD07E2postcritSemRPDlower, VD07E2postcritSemRPDupper),
                   c(VD07E2postcritSemTFTmn, VD07E2postcritSemTFTlower, VD07E2postcritSemTFTupper),

                   c(VD07E2postcritIntFPRTmn, VD07E2postcritIntFPRTlower, VD07E2postcritIntFPRTupper),
                   c(VD07E2postcritIntRPDmn, VD07E2postcritIntRPDlower, VD07E2postcritIntRPDupper),
                   c(VD07E2postcritIntTFTmn, VD07E2postcritIntTFTlower, VD07E2postcritIntTFTupper),

                   c(VD07E2finalSynFPRTmn, VD07E2finalSynFPRTlower, VD07E2finalSynFPRTupper),
                   c(VD07E2finalSynRPDmn, VD07E2finalSynRPDlower, VD07E2finalSynRPDupper),
                   c(VD07E2finalSynTFTmn, VD07E2finalSynTFTlower, VD07E2finalSynTFTupper),

                   c(VD07E2finalSemFPRTmn, VD07E2finalSemFPRTlower, VD07E2finalSemFPRTupper),
                   c(VD07E2finalSemRPDmn, VD07E2finalSemRPDlower, VD07E2finalSemRPDupper),
                   c(VD07E2finalSemTFTmn, VD07E2finalSemTFTlower, VD07E2finalSemTFTupper),

                   c(VD07E2finalIntFPRTmn, VD07E2finalIntFPRTlower, VD07E2finalIntFPRTupper),
                   c(VD07E2finalIntRPDmn, VD07E2finalIntRPDlower, VD07E2finalIntRPDupper),
                   c(VD07E2finalIntTFTmn, VD07E2finalIntTFTlower, VD07E2finalIntTFTupper),

                   c(VD07E3precritSynFPRTmn, VD07E3precritSynFPRTlower, VD07E3precritSynFPRTupper),
                   c(VD07E3precritSynRPDmn, VD07E3precritSynRPDlower, VD07E3precritSynRPDupper),
                   c(VD07E3precritSynTFTmn, VD07E3precritSynTFTlower, VD07E3precritSynTFTupper),

                   c(VD07E3precritSemFPRTmn, VD07E3precritSemFPRTlower, VD07E3precritSemFPRTupper),
                   c(VD07E3precritSemRPDmn, VD07E3precritSemRPDlower, VD07E3precritSemRPDupper),
                   c(VD07E3precritSemTFTmn, VD07E3precritSemTFTlower, VD07E3precritSemTFTupper),

                   c(VD07E3precritIntFPRTmn, VD07E3precritIntFPRTlower, VD07E3precritIntFPRTupper),
                   c(VD07E3precritIntRPDmn, VD07E3precritIntRPDlower, VD07E3precritIntRPDupper),
                   c(VD07E3precritIntTFTmn, VD07E3precritIntTFTlower, VD07E3precritIntTFTupper),

                   c(VD07E3critSynFPRTmn, VD07E3critSynFPRTlower, VD07E3critSynFPRTupper),
                   c(VD07E3critSynRPDmn, VD07E3critSynRPDlower, VD07E3critSynRPDupper),
                   c(VD07E3critSynTFTmn, VD07E3critSynTFTlower, VD07E3critSynTFTupper),

                   c(VD07E3critSemFPRTmn, VD07E3critSemFPRTlower, VD07E3critSemFPRTupper),
                   c(VD07E3critSemRPDmn, VD07E3critSemRPDlower, VD07E3critSemRPDupper),
                   c(VD07E3critSemTFTmn, VD07E3critSemTFTlower, VD07E3critSemTFTupper),

                   c(VD07E3critIntFPRTmn, VD07E3critIntFPRTlower, VD07E3critIntFPRTupper),
                   c(VD07E3critIntRPDmn, VD07E3critIntRPDlower, VD07E3critIntRPDupper),
                   c(VD07E3critIntTFTmn, VD07E3critIntTFTlower, VD07E3critIntTFTupper),

                   c(VD07E3postcritSynFPRTmn, VD07E3postcritSynFPRTlower, VD07E3postcritSynFPRTupper),
                   c(VD07E3postcritSynRPDmn, VD07E3postcritSynRPDlower, VD07E3postcritSynRPDupper),
                   c(VD07E3postcritSynTFTmn, VD07E3postcritSynTFTlower, VD07E3postcritSynTFTupper),

                   c(VD07E3postcritSemFPRTmn, VD07E3postcritSemFPRTlower, VD07E3postcritSemFPRTupper),
                   c(VD07E3postcritSemRPDmn, VD07E3postcritSemRPDlower, VD07E3postcritSemRPDupper),
                   c(VD07E3postcritSemTFTmn, VD07E3postcritSemTFTlower, VD07E3postcritSemTFTupper),

                   c(VD07E3postcritIntFPRTmn, VD07E3postcritIntFPRTlower, VD07E3postcritIntFPRTupper),
                   c(VD07E3postcritIntRPDmn, VD07E3postcritIntRPDlower, VD07E3postcritIntRPDupper),
                   c(VD07E3postcritIntTFTmn, VD07E3postcritIntTFTlower, VD07E3postcritIntTFTupper),

                   c(VD07E3finalSynFPRTmn, VD07E3finalSynFPRTlower, VD07E3finalSynFPRTupper),
                   c(VD07E3finalSynRPDmn, VD07E3finalSynRPDlower, VD07E3finalSynRPDupper),
                   c(VD07E3finalSynTFTmn, VD07E3finalSynTFTlower, VD07E3finalSynTFTupper),

                   c(VD07E3finalSemFPRTmn, VD07E3finalSemFPRTlower, VD07E3finalSemFPRTupper),
                   c(VD07E3finalSemRPDmn, VD07E3finalSemRPDlower, VD07E3finalSemRPDupper),
                   c(VD07E3finalSemTFTmn, VD07E3finalSemTFTlower, VD07E3finalSemTFTupper),

                   c(VD07E3finalIntFPRTmn, VD07E3finalIntFPRTlower, VD07E3finalIntFPRTupper),
                   c(VD07E3finalIntRPDmn, VD07E3finalIntRPDlower, VD07E3finalIntRPDupper),
                   c(VD07E3finalIntTFTmn, VD07E3finalIntTFTlower, VD07E3finalIntTFTupper)
            )

VD07expt <- c(rep("VD07.E1 (N=35)", 9), rep("VD07.E2 (N=36)", 27), rep("VD07.E3 (N=40)", 36))

VD07region <- c(rep("critical", 3), rep("spill-over", 3), rep("sentence-final", 3),
                rep("critical", 9), rep("spill-over", 9), rep("sentence-final", 9),
                rep("pre-critical", 9), rep("critical", 9), rep("spill-over", 9), rep("sentence-final", 9)
                )

VD07int_type <- c(rep(c("syntactic", "semantic", "interaction"), 3),

                  rep(c("syntactic", "syntactic", "syntactic", 
                        "semantic", "semantic", "semantic",
                        "interaction", "interaction", "interaction"), 7)
                  )

VD07measure <- c(rep("SPR", 9), rep(c("FPRT", "RPD","TFT"), 21))

VD07estimates <- as.data.frame(cbind(VD07summary, 
                                     VD07expt,
                                     VD07region,
                                     VD07int_type,
                                     VD07measure))

colnames(VD07estimates)<-c("mean", "lower" ,"upper",  "experiment", "region", "int_type","measure")

VD07estimates$mean <- as.numeric(VD07estimates$mean)
VD07estimates$lower <- as.numeric(VD07estimates$lower)
VD07estimates$upper <- as.numeric(VD07estimates$upper)

VD07estimates$experiment <- factor(VD07estimates$experiment,levels=c("VD07.E1 (N=35)", "VD07.E2 (N=36)","VD07.E3 (N=40)"))

VD07estimates$measure <- factor(VD07estimates$measure,levels=c("SPR","FPRT","RPD","TFT"))

VD07estimates$region <- factor(VD07estimates$region,levels=c("pre-critical","critical", "spill-over", "sentence-final"))

VD07estimates$int_type <- factor(VD07estimates$int_type,levels=c("syntactic","semantic", "interaction"))
```

# Mertzen et al 2023: 
```{r}
load("M21_byregionsummaries_de.rda")
M23.G <- d
M23.G$experiment <- "M23.G (N=121)"

load("M21_byregionsummaries_en.rda")
M23.E <- d
M23.E$experiment <- "M23.E (N=61)"

M23estimates <- rbind(M23.G, M23.E[,-7])

M23estimates$int_type <- ifelse(M23estimates$effect%in%c("Subject"),"syntactic",
                              ifelse(M23estimates$effect%in%c("Animacy"),"semantic", 
                                   "interaction"))

M23estimates$region <- as.factor(M23estimates$region)
levels(M23estimates$region)[3] <- "spill-over"

M23estimates <- M23estimates %>% select(-lang, -region.label, -effect)
```

# Plot present and previous findings
```{r}
my_colors <- RColorBrewer::brewer.pal(11, "RdBu")

# prepare Schoknecht & Vasishth estimates (numbers from script "Plot_stats.R")
SVestimates <- data.frame(mean   = c(3, 14, 8, -2, 10, 14, 8, 14, 8),
                          lower  = c(-4, 7, -5, -11, 1, -3, 1, 7, -6),
                          upper = c(10, 21, 22, 7, 19, 31, 15, 21, 21),
                           region = c(rep(c("pre-critical"),3), rep(c("critical"), 3), rep(c("spill-over"), 3)),
                           measure  = c(rep(c("SPR"), 9)),
                           experiment = c(rep(c("present (N=774)"), 9)),
                           int_type = c(rep(c("syntactic", "semantic", "interaction"), 3))
                          )

# combine with previous estimates
estimates <- rbind(VD07estimates, M23estimates, SVestimates)
estimates$experiment <- factor(estimates$experiment, levels=c("VD07.E1 (N=35)", "VD07.E2 (N=36)", "VD07.E3 (N=40)", "M23.E (N=61)", "M23.G (N=121)", "present (N=774)"))

RT_plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over")),
                       aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
                  geom_pointrange(aes(color=experiment, shape=measure), 
                       size=0.8, position=position_dodge(1.45))+
                  scale_shape_manual(values = c(18,16,17,15), breaks=c("SPR", "FPRT", "RPD", "TFT"),
                                     labels=c("self-paced reading time", "first-pass reading time", "regression path duration", "total fixation time"))+
                  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
                  geom_hline(yintercept=0, linetype="solid", alpha=.5)+
                  facet_grid(region~int_type, scales="free", space="free") +
                  theme_bw(base_size = 15)+
                  scale_x_discrete(breaks = NULL)+
                  theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
                  xlab("")+
                  ylab("reading time difference (ms)")+
                  #ggtitle("Present and previous findings")+
                  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
                  coord_cartesian(ylim=c(-75, 150)) +
                  ggtitle("Reading time differences") 
ggsave("plots/present_previous_findings.jpg", width=9, height=8, dpi=600)

TFT_plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over") & measure %in%c("SPR", "TFT")), aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
                  geom_pointrange(aes(color=experiment, shape=measure), 
                           size=0.8, position=position_dodge(1.45))+
                scale_shape_manual(values = c(18,16,17,15), breaks=c("SPR", "FPRT", "RPD", "TFT"),
                                       labels=c("self-paced reading time", "first-pass reading time", "regression path duration", "total fixation time"))+
                scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
                geom_hline(yintercept=0, linetype="solid", alpha=.5)+
                facet_grid(region~int_type, scales="free", space="free") +
                theme_bw(base_size = 15)+
                scale_x_discrete(breaks = NULL)+
                theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
                xlab("")+
                ylab("reading time difference (ms)")+
                #ggtitle("Present and previous findings")+
                guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
                coord_cartesian(ylim=c(-50, 100))  
ggsave("plots/present_previous_findings_SPR_TFT.jpg", width=9, height=8, dpi=600)

```

# Compute BF10 for Van Dyke (2007) data
```{r}
# function to compute BF10 repeatedly
compute_BF <- function(mean_effect, se){

  #universal values for a reasonable prior: N(0,0.5), on the ms scale [-40, 40]
    xbar <- mean_effect
    SE <- se
    mu_prior<-0
    prior_variance<-40^2
    a <- 1/prior_variance
    b <- 1/SE^2
    mu_posterior <- (a*mu_prior+b*xbar)/(a+b)
    mu_posterior

    var_posterior <- 1/(a+b)
    sqrt(var_posterior)

    mu_posterior <- (a*mu_prior+b*mean_effect)/(a+1/SE^2)
    var_posterior <- 1/(a+(1/SE^2))
    # BF10 (1/(density under posterior at 0 / density under prior at 0):
    1/(dnorm(0,mu_posterior,sqrt(var_posterior))/dnorm(0,0,40))
}

##### E1 critical ####
bf.E1.syn.crit <- compute_BF(VD07E1critSynmn, VD07E1critSynSE)
bf.E1.sem.crit <- compute_BF(VD07E1critSemmn, VD07E1critSemSE)
bf.E1.int.crit <- compute_BF(VD07E1critIntmn, VD07E1critIntSE)

##### E1 spill-over ####
bf.E1.syn.spill <- compute_BF(VD07E1spillSynmn, VD07E1spillSynSE)
bf.E1.sem.spill <- compute_BF(VD07E1spillSemmn, VD07E1spillSemSE)
bf.E1.int.spill <- compute_BF(VD07E1spillIntmn, VD07E1spillIntSE)

##### E2 critical #### 
# for the eye-tracking experiments, we just compute BF10 for TFT 
bf.E2.syn.crit <- compute_BF(VD07E2critSynTFTmn, VD07E2critSynTFTSE)
bf.E2.sem.crit <- compute_BF(VD07E2critSemTFTmn, VD07E2critSemTFTSE)
bf.E2.int.crit <- compute_BF(VD07E2critIntTFTmn, VD07E2critIntTFTSE)

##### E2 spill-over #### 
bf.E2.syn.spill <- compute_BF(VD07E2postcritSynTFTmn, VD07E2postcritSynTFTSE)
bf.E2.sem.spill <- compute_BF(VD07E2postcritSemTFTmn, VD07E2postcritSemTFTSE)
bf.E2.int.spill <- compute_BF(VD07E2postcritIntTFTmn, VD07E2postcritIntTFTSE)

##### E3 pre-critical #### 
bf.E3.syn.precrit <- compute_BF(VD07E3precritSynTFTmn, VD07E3precritSynTFTSE)
bf.E3.sem.precrit <- compute_BF(VD07E3precritSemTFTmn, VD07E3precritSemTFTSE)
bf.E3.int.precrit <- compute_BF(VD07E3precritIntTFTmn, VD07E3precritIntTFTSE)

##### E3 critical #### 
bf.E3.syn.crit <- compute_BF(VD07E3critSynTFTmn, VD07E3critSynTFTSE)
bf.E3.sem.crit <- compute_BF(VD07E3critSemTFTmn, VD07E3critSemTFTSE)
bf.E3.int.crit <- compute_BF(VD07E3critIntTFTmn, VD07E3critIntTFTSE)

##### E3 spill-over #### 
bf.E3.syn.spill <- compute_BF(VD07E3postcritSynTFTmn, VD07E3postcritSynTFTSE)
bf.E3.sem.spill <- compute_BF(VD07E3postcritSemTFTmn, VD07E3postcritSemTFTSE)
bf.E3.int.spill <- compute_BF(VD07E3postcritIntTFTmn, VD07E3postcritIntTFTSE)

# save BFs in df
df.bf_vd <- data.frame(matrix(ncol=8,nrow=0))
colnames(df.bf_vd) <- c("Experiment", "BF_method", "Language", "Region", "Effect","Prior", "truncated", "BF10")

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E1.syn.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E1.sem.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E1.int.crit,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E1.syn.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E1.sem.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E1.int.spill,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E2.syn.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E2.sem.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E2.int.crit,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E2.syn.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E2.sem.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E2.int.spill,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.precrit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.precrit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.precrit,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.crit,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.crit,2))

df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.spill,2))
df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.spill,2))
```

# Plot BFs
```{r}
# load other BFs
load("BFs_mertzen23.Rda")
df.bf_m23$Experiment <- ifelse(df.bf_m23$Language == "English", "M23.E", "M23.G")

load("BFs_schoknecht.Rda")
df.bf_schoknecht$Experiment <- "present"

df.bf <- rbind(df.bf_vd, df.bf_m23, df.bf_schoknecht)

# prepare
df.bf$Prior2 <- ifelse(df.bf$Prior == "Normal(0, 0.01)", "N(0, 0.01)", 
                        ifelse(df.bf$Prior == "Normal(0, 0.05)", "N(0, 0.05)", 
                              ("N(0, 0.1)")))
df.bf$BF10 <- as.numeric(df.bf$BF10)
df.bf$effect <- factor(df.bf$Effect, levels=c("syntactic","semantic", "interaction"))
df.bf$region <- factor(df.bf$Region, levels=c("pre-critical", "critical", "spill-over"))
df.bf$experiment <- factor(df.bf$Experiment, levels=c("VD07.E1", "VD07.E2", "VD07.E3", "M23.E", "M23.G", "present"))

bfs_plot_1prior <- ggplot(data=filter(df.bf, Prior2 == "N(0, 0.05)"),
                                 aes(x=experiment, y=BF10)) +
                      geom_line(aes(),color="grey", group=1, linewidth=1) +
                    geom_point(aes(color=experiment, fill=experiment, shape=experiment), size=4) +
                    scale_shape_manual(values = c(18,15,15,15,15,18))+
                scale_fill_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
                scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
                geom_hline(yintercept=1, linetype="dashed")+
                facet_grid(region~effect, scales="free", space="free") +
                theme_bw(base_size = 15)+
                scale_x_discrete(breaks = NULL)+
                theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
                xlab("")+
              scale_y_log10("Bayes factor (BF10)",
                             breaks =  c(0.05, 0.1, 0.5, 1, 3, 10, 50, 100, 200, 500, 75000),
                             labels = c(0.05, 0.1, 0.5, 1,  3, 10, 50, 100, 200, 500, 75000)
                            )+
                guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
                ggtitle("Bayes factors under prior N(0, 0.05)")
ggsave("plots/present_previous_BFs_1Prior.jpg", width=9, height=8, dpi=600)

ggplot(data=df.bf,aes(x=Prior2, y=BF10, group=experiment)) +
      geom_line(aes(color=experiment)) +
      geom_point(aes(color=experiment), size=2) +
  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
  geom_hline(yintercept=1, linetype="dashed")+
  facet_grid(region~effect, scales="free", space="free") +
  theme_bw(base_size = 15)+
  #scale_x_discrete(breaks = NULL)+
  theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
  xlab("")+
scale_y_log10("Bayes factor (BF10)",
               breaks =  c(0.05, 0.1, 0.5, 1, 3, 10, 50, 100, 200, 500, 75000),
               labels = c(0.05, 0.1, 0.5, 1,  3, 10, 50, 100, 200, 500, 75000)
              )+
  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
ggsave("plots/present_previous_BFs_allPriors.jpg", width=11, height=8, dpi=600)

# combined plot
plot_grid(RT_plot, bfs_plot_1prior + theme(legend.position = "none"), rel_widths =c(1,0.6), nrow=1, labels=c("A", "B"))
ggsave("plots/present_previous_allplots.jpg", width=15, height=8, dpi=600)

```